SST_patterns

### 2020-04-21 SST_patterns.Rmd
edits

# # ### The temperature products
# load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
# load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
# load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
# load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
# load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
```

@@

# # All the code below can be viewed in upwell_IDX.Rmd
# # Next to run upwelling index
# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
# load("Data/UI_angle.RData")
# 
upwelling <- UI_angle %>%
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t),
  #        max_t = max(t)) %>%
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>%
  select(-data, -clim) %>%
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

>>>>

# ### The temperature products
# load("~/Documents/Upwelling_MCS/Data/Final_G1SST.RData") #G1SST
load("~/Documents/Upwelling_MCS/Data/Final.RData") # MUR
load("~/Documents/Upwelling_MCS/Data/SACTN_US.RData") # SACTN
load("~/Documents/Upwelling_MCS/Data/OISST_fill.RData") # OISST
load("~/Documents/Upwelling_MCS/Data/CMC_fill.RData") # CMC temperature
```

@@

# # All the code below can be viewed in upwell_IDX.Rmd
# # Next to run upwelling index
# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
# load("Data/UI_angle.RData")
# 
upwelling <- UI_angle %>%
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t),
  #        max_t = max(t)) %>%
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2010-06-23", "2014-08-15")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>%
  select(-data, -clim) %>%
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

# Now applying the Upwelling func
# NB: This only pulls out the event results and not the climatology results
# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
# To create a list output one would use group_map(),
# but this then loses the labels of which sites etc. the results belong to
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event # 1 or 3?
  return(res)
}
# 
ts2clm_custom <- function(df){
  # The climatology base period used here is up for debate...
  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2010-06-23", "2014-08-15"))
  return(res)
}
# 
# Calculate the upwelling event metrics
upwelling_detect_event <- function(df){
  upwell_base <- df %>%
    dplyr::rename(t = date) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~ts2clm_custom(.x)) %>%
    left_join(upwelling, by = c("site", "t")) %>%
    filter(!is.na(exceedance)) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~detect_event_custom(.x))
  }
# 
OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
MUR_upwell_base <- upwelling_detect_event(df = Final)

# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
SACTN_US <- SACTN_US %>%
  filter(site %in% selected_sites)
